---
title: "ChlamydiaScreeningModel_9.28.22"
output: html_document
date: '2022-09-028'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Setting up working environment
```{r}
library(EpiModel)
library(network)

```

##OVERALL GOAL:
#Creating a network model for a US high school
#Using data Admiraal and Handcock: https://projecteuclid-org.proxy.library.upenn.edu/journals/annals-of-applied-statistics/volume-10/issue-4/Modeling-concurrency-and-selective-mixing-in-heterosexual-partnership-networks-with/10.1214/16-AOAS963.full


#Most basic model - population of 2000 with no race/sex specific parameters
```{r}
set.seed(3972)

##BASIC MODEL - no specification of sex or race
#Assume population of 2000

num.g1 <- num.g2 <- 1000
nw <- network_initialize(n = num.g1 + num.g2)
nw

##Degree Distributions - Population size of 2000
#Degree     Black Males   White Males   Black Females   White Females
#0              80            320           80              320
#1              100           400           100             400
#2              12            48            12              48
#3              6             24            6               24
#4              2             8             2               8
#Pop Size       200           800           200             800

#Mean degree = 0.75

##SIMPLE MODEL - only includes specification of number of degree 0 individuals and number of concurrent partnerships. No specification of sex or racial mixing.
formation <- ~edges + degree(0) + concurrent

#Edges target stat = 2000*0.75/2 = 750

target.stats <- c(750, 800, 200)

#Dissolution - paper says average relationship is 4 months, scale models in days
coef.diss <- dissolution_coefs(dissolution = ~offset(edges), duration = 120)

est <- netest(nw, formation, target.stats, coef.diss)

#Network diagnostics - 10 simulations, 4 years (1460 days)
dx <- netdx(est, nsims = 10, nsteps = 1460, ncores = 16, nwstats.formula = ~edges + degree(0) + concurrent)
dx
plot(dx)

```


#Slightly more complicated model - sex-specific parameters included, assuming only white adolescents using paper's reported degree distributions
```{r}
set.seed(3972)

#Assume population of 2000, 50% males 50% females
num1 <- num2 <- 1000
n <- 2000
nw <- network_initialize(n = num1 + num2)
nw <- set_vertex_attribute(nw, "group", rep(1:2, c(num1, num2)))

statusVec <- rep("s", n)
init.latent <- sample(1:n, 50)
statusVec[init.latent] <- "e"

statusTime <- rep(NA, n)
statusTime[which(statusVec == "e")] <- 1

symptomatic <- rep(NA, n)

nw <- set_vertex_attribute(nw, "status", statusVec)
nw <- set_vertex_attribute(nw, "statusTime", statusTime)
nw <- set_vertex_attribute(nw, "symptomatic", symptomatic)
nw

##Degree Distributions - Population size of 2000
#Degree        Males        Females
#0              382           345
#1              566           630
#2              40            23
#3              12            2
#Pop Size       682           682

#Mean degree = 0.682

#Setting degree distributions for female (g1) and males (g2). Using degree 0-3 based on reported relationships resulted in unbalanced edges. Tweaked female degree 0, 1, and 2 to end up with equal edges between males and females.
deg.dist.g1 <- c(0.345, 0.63, 0.023, 0.002)
deg.dist.g2 <- c(0.382, 0.566, 0.040, 0.012)

check_degdist_bal(num1, num2, deg.dist.g1, deg.dist.g2)


##Formation of model
formation <- ~edges + degree(0:3, by="group") + nodematch("group")

#Edges target stat = 2000*0.682/2 = 682

target.stats <- c(682, 345, 630, 23, 2, 382, 566, 40, 12, 0)

#Dissolution - paper says average relationship is 4 months, scale models in days
coef.diss <- dissolution_coefs(dissolution = ~offset(edges), duration = 120)

est <- netest(nw, formation, target.stats, coef.diss)

dx <- netdx(est, nsims = 10, nsteps = 1460, ncores = 16, nwstats.formula = ~ edges + degree(0:3, by="group") + nodematch("group"))
dx
plot(dx)

```

#Setting disease parameters
#Davies et al. Theoretical Biology and Medical Modelling 2014, 11:8
```{r}
#
param <- param.net(inf.prob = 0.095, inf.prob.g2 = 0.095, act.rate = 2, rec.rate = 0.1, rec.rate.g2 = 0.1)

#10 infected males, 10 infected females
init <- init.net(i.num = 10, i.num.g2 = 10)

#SIS model lasting 1460 days, 10 simulations
control <- control.net(type = "SIS", nsims = 10, nsteps = 1460)

sim <- netsim(est, param, init, control)
sim

plot(sim, type = "network", col.status = TRUE, at = 1, sims = 1)

#Run one simulation at a time

```


#Customizing base SEIR model to include additional compartments
```{r}
##Customizing compartment flow


#S = Susceptible
#E = Exposed, not infectious
#Ia = Infected, asymptomatic
#Is = Infected, symptomatic
#R = Recovered (temporarily immune)

progress2 <- function(dat, at) {

  ## Attributes
  active <- get_attr(dat, "active")
  status <- get_attr(dat, "status")
  statusTime <- get_attr(dat, "statusTime")
  symptomatic <- get_attr(dat, "symptomatic")

  ## Parameters
  prop.sympt <- get_param(dat, "prop.sympt")
  eis.rate <- get_param(dat, "eis.rate")
  iss.rate <- get_param(dat, "iss.rate")
  eia.rate <- get_param(dat, "eia.rate")
  iar.rate <- get_param(dat, "iar.rate")
  rs.rate <- get_param(dat, "rs.rate")
  ias.rate <- get_param(dat, "ias.rate")

  ## Determine Symptomatic (E to Is) or Asymptomatic (E to Ia) pathway
  ids.newInf <- which(active == 1 & status == "E" & statusTime <= at & is.na(symptomatic))
  num.newInf <- length(ids.newInf)
  if (num.newInf > 0) {
    prop.sympt.vec <- prop.sympt
    vec.new.sympt <- rbinom(num.newInf, 1, prop.sympt.vec)
    sympt[ids.newInf] <- vec.new.sympt
  }

  
  ## Asymptomatic Pathway
  # E to Ia: latent move to asymptomatic infectious
  num.new.EtoIa <- 0
  ids.Eia <- which(active == 1 & status == "E" & statusTime < at & symptomatic == 0)
  num.Eia <- length(ids.Eia)
  if (num.Eia > 0) {
    vec.new.Ia <- which(rbinom(num.Eia, 1, eia.rate) == 1)
    if (length(vec.new.Ia) > 0) {
      ids.new.Ia <- ids.Eia[vec.new.Ia]
      num.new.EtoIa <- length(ids.new.Ia)
      status[ids.new.Ia] <- "Ia"
      statusTime[ids.new.Ia] <- at
    }
  }

  # Ia to R: asymptomatic infectious move to recovered (temporarily immune)
  num.new.IatoR <- 0
  ids.Ia <- which(active == 1 & status == "Ia" & statusTime < at & symptomatic == 0)
  num.Ia <- length(ids.Ia)
  if (num.Ia > 0) {
    vec.new.R <- which(rbinom(num.Ia, 1, iar.rate) == 1)
    if (length(vec.new.R) > 0) {
      ids.new.R <- ids.Ia[vec.new.R]
      num.new.IatoR <- length(ids.new.R)
      status[ids.new.R] <- "R"
      statusTime[ids.new.R] <- at
    }
  }
  
  # R to S: recovered (temporarily immune) to susceptible
  num.new.RtoS <- 0
  ids.R <- which(active == 1 & status == "R" & statusTime < at & symptomatic == 0)
  num.R <- length(ids.R)
  if (num.R > 0) {
    vec.new.S <- which(rbinom(num.R, 1, rs.rate) == 1)
    if (length(vec.new.R) > 0) {
      ids.new.S <- ids.R[vec.new.S]
      num.new.RtoS <- length(ids.new.S)
      status[ids.new.S] <- "S"
      statusTime[ids.new.S] <- at
    }
  }
  
    # Ia to S: Asymptomatic to susceptible (identified through screening)
  num.new.IatoS <- 0
  ids.Ia <- which(active == 1 & status == "Ia" & statusTime < at & symptomatic == 0)
  num.Ia <- length(ids.Ia)
  if (num.Ia > 0) {
    vec.new.S <- which(rbinom(num.Ia, 1, ias.rate) == 1)
    if (length(vec.new.R) > 0) {
      ids.new.S <- ids.R[vec.new.S]
      num.new.RtoS <- length(ids.new.S)
      status[ids.new.S] <- "S"
      statusTime[ids.new.S] <- at
    }
  }

  ## Symptomatic Pathway
  # E to Is: latent move to symptomatic infectious
  num.new.EtoIs <- 0
  ids.Eis <- which(active == 1 & status == "E" & statusTime < at & symptomatic == 1)
  num.Eis <- length(ids.Eis)
  if (num.Eis > 0) {
    vec.new.Is <- which(rbinom(num.Eis, 1, eis.rate) == 1)
    if (length(vec.new.Is) > 0) {
      ids.new.Is <- ids.Eis[vec.new.Is]
      num.new.EtoIs <- length(ids.new.Is)
      status[ids.new.Is] <- "Is"
      statusTime[ids.new.Is] <- at
    }
  }

  # Is to S: Symptomatic infectious move to susceptible
  num.new.IstoS <- 0
  ids.Is <- which(active == 1 & status == "Is" & statusTime < at & symptomatic == 1)
  num.Is <- length(ids.Is)
  if (num.Is > 0) {
    vec.new.Is <- which(rbinom(num.Is, 1, iss.rate) == 1)
    if (length(vec.new.S) > 0) {
      ids.new.S <- ids.Is[vec.new.S]
      num.new.IstoS <- length(ids.new.S)
      status[ids.new.S] <- "S"
      statusTime[ids.new.S] <- at
    }
  }

  
  ## Save updated status attribute
  dat <- set_attr(dat, "status", status)
  dat <- set_attr(dat, "statusTime", statusTime)
  dat <- set_attr(dat, "symptomatic", symptomatic)

  ## Save summary statistics
  dat <- set_epi(dat, "eis.flow", at, num.new.EtoIs)
  dat <- set_epi(dat, "iss.flow", at, num.new.IstoS)
  dat <- set_epi(dat, "eia.flow", at, num.new.EtoIa)
  dat <- set_epi(dat, "iar.flow", at, num.new.IatoR)
  dat <- set_epi(dat, "rs.flow", at, num.new.RtoS)
  dat <- set_epi(dat, "ias.flow", at, num.new.IatoS)

  dat <- set_epi(dat, "e.num", at, sum(status == "e"))
  dat <- set_epi(dat, "Is.num", at, sum(status == "is"))
  dat <- set_epi(dat, "s.num", at, sum(status == "s"))
  dat <- set_epi(dat, "Ia.num", at, sum(status == "ia"))
  dat <- set_epi(dat, "r.num", at, sum(status == "r"))

  return(dat)
}
  

##Customizing infection module
infect2 <- function(dat, at) {

  ## Attributes ##
  active <- get_attr(dat, "active")
  status <- get_attr(dat, "status")
  infTime <- get_attr(dat, "infTime")
  statusTime <- get_attr(dat, "statusTime")

  ## Parameters ##
  inf.prob <- get_param(dat, "inf.prob")
  act.rate <- get_param(dat, "act.rate")
  inf.prob.sympt.rr <- get_param(dat, "inf.prob.sympt.rr")
  act.rate.dx.sympt.time <- get_param(dat, "act.rate.dx.sympt.time")
  act.rate.dx.sympt.rr <- get_param(dat, "act.rate.dx.sympt.rr")

  ## Find infected nodes ##
  infstat <- c("Ia", "Is")
  idsInf <- which(active == 1 & status %in% infstat)
  nActive <- sum(active == 1)
  nElig <- length(idsInf)

  ## Initialize default incidence at 0 ##
  nInf <- 0

  ## If any infected nodes, proceed with transmission ##
  if (nElig > 0 && nElig < nActive) {

    ## Look up discordant edgelist ##
    del <- discord_edgelist(dat, at, infstat = infstat)

    ## If any discordant pairs, proceed ##
    if (!(is.null(del))) {

      del$status <- status[del$inf]

      # Set parameters on discordant edgelist data frame
      del$transProb <- inf.prob
      del$transProb[del$status == "Is"] <- del$transProb[del$status == "Is"] *
                                          inf.prob.sympt.rr
      del$actRate <- act.rate
      if (at >= act.rate.dx.sympt.time) {
        del$actRate[del$dxStatus == 2] <- del$actRate[del$dxStatus == 2] *
                                          act.rate.dx.sympt.rr
      }
      del$finalProb <- 1 - (1 - del$transProb)^del$actRate

      # Stochastic transmission process
      transmit <- rbinom(nrow(del), 1, del$finalProb)

      # Keep rows where transmission occurred
      del <- del[which(transmit == 1), ]

      # Look up new ids if any transmissions occurred
      idsNewInf <- unique(del$sus)
      nInf <- length(idsNewInf)

      # Set new attributes and transmission matrix
      if (nInf > 0) {
        status[idsNewInf] <- "e"
        infTime[idsNewInf] <- at
        statusTime[idsNewInf] <- at
        dat <- set_transmat(dat, del, at)
      }
    }
  }

  dat <- set_attr(dat, "status", status)
  dat <- set_attr(dat, "infTime", infTime)
  dat <- set_attr(dat, "statusTime", statusTime)

  ## Save summary statistics
  dat <- set_epi(dat, "se.flow", at, nInf)

  return(dat)
}


```

#Setting infection parameters
```{r}
param <- param.net(inf.prob = 0.1,
                   act.rate = 3,
                   inf.prob.sympt.rr = 0.5,
                   act.rate.dx.sympt.time = Inf,
                   act.rate.dx.sympt.rr = 0.05,
                   eis.rate = 1/4.0,
                   iss.rate = 1/5.0,
                   eia.rate = 1/4.0,
                   iar.rate = 1/1.5,
                   rs.rate = 1/3.5,
                   ias.rate = 1/2.5,
                   prop.sympt = 0.85)
init <- init.net(i.num = 10, i.num.g2 = 10)

```

```{r}
control <- control.net(type = NULL,
                       nsims = 10,
                       ncores = 5,
                       nsteps = 100,
                       infection.FUN = infect2,
                       progress.FUN = progress2,
                       resimulate.network = TRUE,
                       tergmLite = TRUE,
                       set.control.tergm = 
                         control.simulate.formula.tergm(MCMC.burnin.min = 10000))

sim <- netsim(est, param, init, control)
sim


plot(sim)

```




#Random disease parameters just to plot network
#param <- param.net(inf.prob = 0.4, inf.prob.g2 = 0.2, act.rate = 2, rec.rate = 0.1, rec.rate.g2 = 0.1)
#init <- init.net(i.num = 10, i.num.g2 = 10)
#control <- control.net(type = "SIS", nsims = 5, nsteps = 1460)

#sim <- netsim(est, param, init, control)

#plot(sim, type = "network", col.status = TRUE, at = 1, sims = 1)

```

